---
title: "UNHCR | Survey Designer"
output: 
  flexdashboard::flex_dashboard:
    orientation: column
    source_code: embed
    theme: 
      version: 4
      navbar-bg: "#0072BC"
      bootswatch: yeti
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)

library(shinyBS) ## For tooltip

library(jsonlite)
library(shinipsum)
library(DT)

 
#devtools::install_github("shinyTree/shinyTree")
#library(shinyTree)

#devtools::install_github("stla/jsTreeR")
library(jsTreeR)  
```


Sidebar {.sidebar}
=====================================

Get Household Survey Questionnaires fully aligned with international standards and internal codebook. 

```{r}
tags$div(title="Here you can select the Contextualised Library of indicators",
selectInput("library", label =  "Regional Library" , 
    choices = list("Americas" = "Americas", 
                   "Asia" = "Asia",
                   "Eastern-Africa" = "Eastern-Africa",
                   "Western-Africa" = "Western-Africa", 
                   "Southern-Africa" = "Southern-Africa", 
                   "MENA" = "MENA",
                   "Europa" = "Europa"), 
    selected = "Americas"
    )
)


```

```{r}
# https://www.davidsolito.com/post/conditional-drop-down-in-shiny/ 
tags$div(title="Select your country - this will pull up automatically the Case Load that should be covered based on your annual statistical report and implement the languages translation and geographic breakdown required for the country",
selectInput("library", 
            label = "Country → PopGroup, Lang & Pcode", 
    choices = list("Panama" = "Panama", 
                   "Colombia" = "Colombia",
                   "Ecuador" = "Ecuador" ), 
    selected = "Panama")
)
```


```{r}
tags$div(title="Define how many data collection waves you can manage within the year. The more data collection wave the more indicators can be collected. ",
selectInput("wave", 
            label = "How many data collection waves for this annual cycle?", 
    choices = list("Four-waves" = "Four waves",
                   "Three-waves" = "Three waves",
                   "One-wave" = "One wave", 
                   "Two-waves" = "Two waves" ), 
    selected = "Four-waves")
)
``` 

```{r}
tags$div(title="Using multiple data collection mode in parallel can help minimizing non-reponse and increase coverage.",
checkboxGroupInput("mode", 
              label =  "What Data Collection Modes can be used for this annual cycle?" , 
    # choiceNames = list("Face-to-face CAPI"  ,
    #                "Phone-Interview CATI"  ,
    #                "Self-administered CAWI"  ),
      
    choiceNames = list(
      list(
        icon("calendar-check-o"),
         "Face-to-face CAPI"
      ), 
      list(
        icon("phone-square") ,
        "Phone-Interview CATI"  
      ), 
      list(
        icon("pencil-square-o"),
        "Self-administered CAWI"  
      )),
    
   choiceValues =  list("Face-to-face"  , 
                   "Phone-Interview"  ,
                   "Self-administered" ))
)
```

 

```{r}
checkboxGroupInput("topic", 
              label = "Select the Topics to prioritise in the annual data collection cycle!", 
    choiceNames = list(
      list(
        icon("gavel") ,
        "Rights"  
      ), 
      list(
        icon("home"),
        "Trajectory"  
      ), 
      list(
        icon("thermometer-full"),
        "Intention"  
      ), 
      list(
        icon("check-circle"),
        "Basic Needs"  
      ), 
      list(
        icon("certificate"),
        "Coping Capacity"  
      ), 
      list(
        icon("users"),
        "Well-Being"  
      ), 
      list(
        icon("blind"),
         "Victimisation"
      ), 
      list(
        icon("usd"),
        "Livelihood"  
      ), 
      list(
        icon("book"),
        "Education"  
      ), 
      list(
        icon("medkit"),
        "Health-Nutrition"  
      ), 
      list(
        icon("eye"),
        "Accountability" 
      ), 
      list(
        icon("info-circle"),
        "Information" 
      )),
   choiceValues =  list("Rights"  ,
                   "Trajectory"  , 
                   "Intention"  ,  
                   "Basic Needs"  , 
                   "Coping Capacity"  , 
                   "Well-Being"  ,
                   "Victimisation"  , 
                   "Livelihood"  ,
                   "Education"  ,
                   "Health-Nutrition"  , 
                   "Accountability",
                   "Information"  ))


 

```


```{r}
nodes <- list(
  list(
    text = "Standard Indicators",
    data = list(value = 999),
    icon = "far fa-file red",
    children = list(
      list(
        text =  "Rights"   ,
        icon = "fa fa-gavel green",
        children = list(
          list(
            text = "SDG Indicator 16.2.1" 
          ),
          
          list(
            text = "SDG Indicator 3.2.1"  
          ))
      ), 
      list(
        text = "Trajectory"  ,
        icon = "fa fa-home green"
      ), 
      list(
        text = "Intention"  ,
        icon = "fa fa-thermometer-full green",
        children = list(
          list(
            text = "SDG Indicator 16.2.1"  
          ),
          
          list(
            text = "SDG Indicator 3.2.1"  
          ))
      ), 
      list(
        text = "Basic Needs"  ,
        icon = "fa fa-check-circle green"
      ), 
      list(
        text = "Coping Capacity"  ,
        icon = "fa fa-certificate green"
      ), 
      list(
        text = "Well-Being"  ,
        icon = "fa fa-users green"
      ), 
      list(
        text = "Victimisation"  ,
        icon = "fa fa-blind green",
        children = list(
          list(
            text = "SDG Indicator 16.2.1"  
          ),
          
          list(
            text = "SDG Indicator 3.2.1"  
          ))
      ), 
      list(
        text = "Livelihood"  ,
        icon = "fa fa-usd green",
        children = list(
          list(
            text = "SDG Indicator 16.2.1"  
          ),
          
          list(
            text = "SDG Indicator 3.2.1"  
          ))
      ), 
      list(
        text = "Education"  ,
        icon = "fa fa-book green",
        children = list(
          list(
            text = "SDG Indicator 16.2.1"  
          ),
          
          list(
            text = "SDG Indicator 3.2.1"  
          ))
      ), 
      list(
        text = "Health-Nutrition"  ,
        icon = "fa fa-medkit green",
        children = list(
          list(
            text = "SDG Indicator 16.2.1"  
          ),
          
          list(
            text = "SDG Indicator 3.2.1"  
          ))
      ), 
      list(
        text = "Accountability" ,
        icon = "fa fa-eye green"
      ), 
      list(
        text = "Information" ,
        icon = "fa fa-info-circle green"
      )
    )
  ),
  list(
    text = "Optional Indicators",
    icon = "far fa-star red",
   
    children = list(
      list(
        text =  "Rights"   ,
        icon = "fa fa-gavel green",
        children = list(
          list(
            text = "SDG Indicator 16.2.1"  
          ),
          
          list(
            text = "SDG Indicator 3.2.1"  
          ))
      ), 
      list(
        text = "Trajectory"  ,
        icon = "fa fa-home green"
      ), 
      list(
        text = "Intention"  ,
        icon = "fa fa-thermometer-full green",
        children = list(
          list(
            text = "SDG Indicator 16.2.1"  
          ),
          
          list(
            text = "SDG Indicator 3.2.1"  
          ))
      ), 
      list(
        text = "Basic Needs"  ,
        icon = "fa fa-check-circle green"
      ), 
      list(
        text = "Coping Capacity"  ,
        icon = "fa fa-certificate green"
      ), 
      list(
        text = "Well-Being"  ,
        icon = "fa fa-users green"
      ), 
      list(
        text = "Victimisation"  ,
        icon = "fa fa-blind green",
        children = list(
          list(
            text = "SDG Indicator 16.2.1"  
          ),
          
          list(
            text = "SDG Indicator 3.2.1"  
          ))
      ), 
      list(
        text = "Livelihood"  ,
        icon = "fa fa-usd green",
        children = list(
          list(
            text = "SDG Indicator 16.2.1"  
          ),
          
          list(
            text = "SDG Indicator 3.2.1"  
          ))
      ), 
      list(
        text = "Education"  ,
        icon = "fa fa-book green",
        children = list(
          list(
            text = "SDG Indicator 16.2.1"  
          ),
          
          list(
            text = "SDG Indicator 3.2.1"  
          ))
      ), 
      list(
        text = "Health-Nutrition"  ,
        icon = "fa fa-medkit green",
        children = list(
          list(
            text = "SDG Indicator 16.2.1"  
          ),
          
          list(
            text = "SDG Indicator 3.2.1"  
          ))
      ), 
      list(
        text = "Accountability" ,
        icon = "fa fa-eye green"
      ), 
      list(
        text = "Information" ,
        icon = "fa fa-info-circle green"
      )
    )
  )
)

## Need to filter the orginal node - based on the checkboxGroupInput

## Selector
output[["jstree"]] <- renderJstree({
  jstree(nodes, 
         dragAndDrop = TRUE, 
         checkboxes = TRUE, 
         search = TRUE,
         theme = "proton")
})

## Results of selection
output[["treeSelected"]] <- renderPrint({
  input[["jstree_selected"]]
})
```

1. Indicators selection
=====================================  
 

Column {data-width=400}
-----------------------------------------------------------------------

### Select additional Standard & Optional <code>Indicators</code>

```{r}
helpText(HTML("Core Indicators, like demographics, and modules like screeening questions are already preselected. Check the availibility of baseline values for the indicators you may consider before selecting them<hr> "))

## need to implement
# if(interactive()){
#   jstreeExample("filtering")
# }

jstreeOutput("jstree")
```

Column {data-width=400}
-----------------------------------------------------------------------

### Selected <code>indicators</code> for your survey plan

```{r}

helpText(HTML("You can export your fully documented survey indicators plan with the button below.  <hr>"))

renderPrint({
    toJSON(input[["jstree_selected"]], 
           pretty = TRUE, 
           auto_unbox = TRUE)
  })

#verbatimTextOutput("treeSelected")

# datatable(as.data.frame("treeSelected"),
#   options = list(paging = TRUE,    ## paginate the output
#                  pageLength = 15,  ## number of rows to output for each page
#                  scrollX = TRUE,   ## enable scrolling on X axis
#                  scrollY = TRUE,   ## enable scrolling on Y axis
#                  autoWidth = TRUE, ## use smart column width handling
#                  server = FALSE,   ## use client-side processing
#                  dom = 'Bfrtip',
#                  buttons = c('excel'), # 'csv', 
#                  columnDefs = list(list(targets = '_all', 
#                                         className = 'dt-center'),
#                                    list(targets = c(0, 8, 9), 
#                                         visible = FALSE))
#                  ),
#   extensions = 'Buttons',
#   selection = 'single', ## enable selection of a single row
#   filter = 'bottom',              ## include column filters at the bottom
#   rownames = FALSE                ## don't show row numbers/names
# )

```
    

   
2. Spread over Data Collection Waves
=====================================     

### Allocated <code>Indicators</code> per wave

```{r}
helpText(HTML("Below is an estimation of Interview lenght per questionnaire  <hr>"))
 
```


3. Revise Resulting Forms
=====================================     

### Chart 2

4. Get the XlsForm
=====================================     

### Download <code>XlsForm</code>  

```{r}
helpText(HTML("The resulting XlsForm can either be adjusted manually or pushed directly to UNHCR Kobo server. You can download the files for each data collection wave  <hr>"))
 
```
    
### Push <code>XlsForm</code>  to the server

```{r}
helpText(HTML("Alternatively push directly to the server  <hr>"))
 
```

